---
title: "re-derive ORR"
author: "wei zou"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(rice)
library(reshape2)
library(plyr)
library(dplyr)
library(knitr)

rice_session(
  {
   # asl <- as.data.frame(  rice_read('root/clinical_studies/RO7092284/CDT30310/BO42592/data_analysis/OS_IA1/qa/outdata_vad/adsl.sas7bdat') )
   # 
    raw <- rice_read_zip('root/clinical_studies/RO7496353/CDT70424/GO44010/source/other/prod/GO44010_Daily_Extract_20240311071607.zip')
   # 
rs <- as.data.frame( rice_read("root/clinical_studies/RO7496353/CDT70424/GO44010/data_analysis/IMC_2nd/data/sdtmv/rs.sas7bdat"))

tu <- as.data.frame( rice_read("root/clinical_studies/RO7496353/CDT70424/GO44010/data_analysis/IMC_2nd/data/sdtmv/tu.sas7bdat"))

tr <- as.data.frame( rice_read("root/clinical_studies/RO7496353/CDT70424/GO44010/data_analysis/IMC_2nd/data/sdtmv/tr.sas7bdat"))

 
},
password = Sys.getenv("ENTIMICE_PASSWORD")
)

outF <- '~/GO44010/results'

rrsp1 <- raw$rrsp1
```


```{r  functions, echo=F}


vector_overlap <- function(v1, v2, removeDup=T){
  
  n1 <- deparse(substitute(v1))
  n2 <- deparse(substitute(v2))
  
  stopifnot(is.vector(v1) && is.vector(v2))
  print(paste('raw input vectors have length', length(v1), 'for',n1,'and', length(v2), 'for',n2))
  if( any(duplicated(v1)) & removeDup ) {
    print(paste('there are', sum(duplicated(v1)),'duplicated entries in',n1,'ignored'))
    v1 <- v1[!duplicated(v1)]
  }
  
  if( any(duplicated(v2)) & removeDup ) {
    print(paste('there are', sum(duplicated(v2)),'duplicated entries in',n2,'ignored'))
    v2 <- v2[!duplicated(v2)]
  }
  
  if( any(is.na(v1))){
    print(paste('there are', sum(is.na(v1)),'entries in',n1,'ingored as missing'))
    v1 <- v1[!is.na(v1)]
  }
  
  if( any(is.na(v2))){
    print(paste('there are', sum(is.na(v2)),'entries in',n2,'ingored as missing'))
    v2 <- v2[!is.na(v2)]
  }
  
  print(paste( length(v1), 'elements from',n1,'and', length(v2), 'elements for',n2,'are compared for overlap below'))
  
  u1 <- unique( union(v1, v2) )
  out1 <- cbind(u1 %in% v1, u1 %in% v2)
  row.names(out1) <- u1 
  colnames(out1) <- c(n1, n2)
  print( table( out1[, 1], out1[, 2], dnn=list(paste('IN', n1 ), paste('IN', n2 )), useNA='ifany')  )
  
  invisible(out1)
  
}


# This function merge ds1 and ds2 by visit variables for regular visits; for unscheduled visit, the 2 data frames are merged by date. The date info in ds1 is a range.  The program will always match 2 records (each from one dataset) with the closest gap


approx_merge <- function(
  ds1 = subset(targetResp, ! visit2 %in% 'SCREENING') , 
  ds2 = subset(rs, RSTESTCD %in% c('TRGRESP') & RSEVAL %in% 'INVESTIGATOR'), 
  ds1.time.var = c('early_date_target','late_date_target'), 
  ds2.time.var='RSDTC', 
  ds1.visit.var = 'visit2',
  ds2.visit.var = 'VISIT',
  ds2.vars =c('RSSTRESC') ){
  
  stopifnot(all(c(ds1.time.var, ds1.visit.var) %in% colnames(ds1)))
  stopifnot(all(c(ds2.time.var, ds2.visit.var) %in% colnames(ds2)))
  stopifnot(all(sapply(ds1.time.var, function(v) class(ds1[[v]])) %in%  'Date'))
  stopifnot(all(sapply(ds2.time.var, function(v) class(ds2[[v]])) %in%  'Date'))
  stopifnot(! any( c(ds2.visit.var, ds2.time.var) %in% colnames(ds1)))
  
  v <- vector_overlap(ds1$USUBJID, ds2$USUBJID)
  common_subj <- row.names(v[ v[,1] & v[,2], ])
  
  dsL <- lapply( common_subj, function(s){
 
    ds1b <- subset(ds1, USUBJID %in% s)
    ds2b <- subset(ds2, USUBJID %in% s)
    
    uns1 <- grepl('UNSCHEDULE', ds1b[[ds1.visit.var]])
    uns2 <- grepl('UNSCHEDULE', ds2b[[ds2.visit.var]])
    
    ds12 <- merge( ds1b[!uns1, ], ds2b[!uns2,  c('USUBJID', ds2.visit.var, ds2.time.var, ds2.vars)], all=T, by.x=c('USUBJID', ds1.visit.var), by.y=c('USUBJID', ds2.visit.var))
    
    ds12[[ds2.visit.var]] <- ds12[[ds1.visit.var]]
    
    ds12$gap_day <- abs(as.integer(ds12[[ds1.time.var[1]]] - ds12[[ds2.time.var]]))
    
    if(!any(c(uns1, uns2))) return (list( ds12 ) )
    
        
     #  if(any(uns1) != any(uns2)) {
     #       
     #    print(ds1b[, c('USUBJID',ds1.visit.var)])
     #    print(ds2b[, c('USUBJID',ds2.visit.var)])
     # #   stop('unscheduled visits are present in only 1 dataset')
     #  }
     #  
      
      
      if(!any(uns1)) return( list( ds12, ds2b[uns2, ] ) ) # so none of unscheduled records from ds2 is matched

      ds1b <- ds1b[uns1, ]
      for (e in c(ds2.visit.var, ds2.vars,'gap_day')) ds1b[, e] <- NA
      for (e in ds2.time.var) ds1b[, e] <- as.Date(NA)
      
      if(!any(uns2)) return(list(rbind(ds12, ds1b))) # add back unscheduled records from ds1 and return

      ds2b <- ds2b[uns2, ]  
      matched <- rep(NA, nrow(ds1b))
      
      for (i in 1:nrow(ds1b)) {
        diff_day <- pmin( abs( as.integer( ds1b[i,ds1.time.var[1]] - ds2b[,ds2.time.var] )),
                          abs( as.integer( ds1b[i,ds1.time.var[2]] - ds2b[,ds2.time.var] )) )
        
        m1 <- which.min(diff_day) 
        
        if(length(m1) != 1){
          print('ds1 records with no close match in ds2: (matched results will show NA for them)')
          print( ds1b[i, c('USUBJID',ds1.visit.var, ds1.time.var)] )
    
        }else{
        
          for (e in c(ds2.time.var, ds2.visit.var, ds2.vars)) ds1b[i, e] <-  ds2b[m1, e]
          
          ds1b[i, 'gap_day'] <- diff_day[m1]
          matched[i] <- m1
        }
      }
      
      
      
      list(rbind(ds12, ds1b), ds2b[-matched[!is.na(matched)],,drop=F] )
    } )
  
  matched <- do.call(rbind, lapply(dsL, function(l) l[[1]]))
  
  stopifnot(  !any(duplicated( apply(matched[, c('USUBJID', ds1.visit.var)], 1, paste, collapse=','))))
  
  list( matched= matched, # original ds1, plus additional variables matched from ds2
  
  leftover_source2 =   do.call(rbind, lapply(dsL, function(l) { # ds2 records what are not matched to any record in ds1
    if(length(l) == 2) {
      l[[2]]
    }else{
      NULL
    }
    })) ,   subject_not_Match =  v[ !(v[,1] & v[,2]),, drop=F] # patients who are not present in both datasets
  )
}


# adjust the visit variable: this is necessary because 'UNSCHEDULED' visits may be mixed together

adjust_visit <- function(d, tlist=targe_list, tid.var='TRLNKID', unschedule_condition = 'VISIT %in% "UNSCHEDULED"', testcd_for_missing_lesion = 'LDIAM') {
  stopifnot(all(c('USUBJID','VISIT') %in% colnames(d)))
  stopifnot(! 'visit2' %in% colnames(d))
  
  # create a visit2 variable that separate multiple visits (eg UNSCHEDULED) that are called the same
  d_b <- ddply(d, .(USUBJID, VISIT), function(d1){
    s <- unique(d1$USUBJID)
    stopifnot(s %in% names(tlist))
      
    d1$visit2 <- d1$VISIT
      
     # simple logic to manipulate the visit name. it may fail easily. we need further check
    # concatenate TRDTC to VISIT if 1) number of rows is larger than the number of non-target lesions or 2) lesion ID duplciates
    if(with(d1, all(eval(parse(text = unschedule_condition))))) {
      if ( nrow(d1) > length(tlist[[s]]) || any(duplicated(d1[[tid.var]])) ) {
        d1$visit2 <- with(d1, paste(VISIT, TRDTC))
        print(paste('concatenating VISIT and DTC for',s,'at', unique(d1$VISIT)))
      }
    }
    
    d1
  })
    
    
  ddply(d_b, .(USUBJID, visit2), function(d1){
      s <- unique(d1$USUBJID)

    if(!all(tlist[[s]] %in% d1[[tid.var]])) { # add records to cover missing target lesions
      print(d1[, c('USUBJID','visit2',tid.var)])
      
      to_add <- setdiff( tlist[[s]], d1[[tid.var]])
      d2 <- data.frame( unique(d1$USUBJID), unique(d1$visit2), to_add, testcd_for_missing_lesion, stringsAsFactors = F)
      colnames(d2) <- c('USUBJID','visit2', tid.var,'TRTESTCD')
      
      d1 <- rbind.fill(d1, d2)
      
      print( paste('adding emtpy rows for', paste(to_add, collapse=',')))
      
    }
      
      # delete duplicated lesions within each visit
      dup1 <- duplicated(d1[[tid.var]])
      if(any(dup1)) {
        print('removing duplicated records')
        print(d1[dup1, c('USUBJID','visit2',tid.var)])
        d1 <- d1[!dup1, ]
      }
      
      # check again; even it reports the algorithm fails, the program does not stop
      
      if( ! ( length(d1[[tid.var]]) == length(tlist[[s]]) && 
             all(sort(d1[[tid.var]]) == sort(tlist[[s]])) ) ) {
        print('the simple algorithm to separate multiple unscheduled visit failed')
        print(d1[,c('USUBJID','visit2',tid.var)])
      }
      
      d1 
    })
           
    

}


# adjust the visit variable for non-target lesions: this is necessary because 'UNSCHEDULED' visits may be mixed together

# a less strict checking on lesion ID here (it seems many non-target lesions are not populated)

adjust_visit_nt <- function(d, tlist=targe_list, tid.var='TRLNKID', unschedule_condition = 'VISIT %in% "UNSCHEDULED"' ) {
  stopifnot(all(c('USUBJID','VISIT') %in% colnames(d)))
  stopifnot(! 'visit2' %in% colnames(d))
  
  # create a visit2 variable that separate multiple visits (eg UNSCHEDULED) that are called the same
  d_b <- ddply(d, .(USUBJID, VISIT), function(d1){
    s <- unique(d1$USUBJID)
    stopifnot(s %in% names(tlist))
      
    d1$visit2 <- d1$VISIT
      
     # simple logic to manipulate the visit name. it may fail easily. we need further check
    # concatenate TRDTC to VISIT if 1) number of rows is larger than the number of non-target lesions or 2) lesion ID duplciates
    if(with(d1, all(eval(parse(text = unschedule_condition))))) {
      if ( nrow(d1) > length(tlist[[s]]) || any(duplicated(d1[[tid.var]])) ) {
        d1$visit2 <- with(d1, paste(VISIT, TRDTC))
        print(paste('concatenating VISIT and DTC for',s,'at', unique(d1$VISIT)))
      }
    }
    
    d1
  })
    
  
  # no data manipulate; just print out unexpected cases  
  ddply(d_b, .(USUBJID, visit2), function(d1){
      s <- unique(d1$USUBJID)

    if(length(tlist[[s]]) != nrow(d1) ) {
        print(paste('the simple algorithm to separate multiple unscheduled visit failed'))
        print(paste('the following NT lesions are expected', paste(tlist[[s]], collapse=',')))
        print(d1[,c('USUBJID','visit2',tid.var)])
      }
      
      d1 
    })
           
    

}


```

#  get lesion IDs


Checks not impleted yet

* target lesion: present or not; 2 max per organ

* non-target lesion: present or not

* new lesions: often remains once appears

```{r, echo=F}
tu1 <- subset(tu, TUEVAL %in% c('', 'INVESTIGATOR') & TUTEST %in% 'Tumor Identification' &    TUORRES %in% c('NON-TARGET','TARGET') ) # target / non target lesions from investigor

tu1_target <- subset(tu1,  TUORRES %in% 'TARGET')
tu1_nt <- subset(tu1,  TUORRES %in% 'NON-TARGET')
# add checks on the target lesions

tu_new <- subset(tu, TUEVAL %in%  c('','INVESTIGATOR') & TUTEST %in% 'Tumor Identification' & TUORRES %in% 'NEW')
# this way of filter gets rid of records from the new lesion assessment form, as their TUORRES will be either 'NEW NON-TARGET' 'NEW TARGET'

tu_new$TUDTC <- as.Date(tu_new$TUDTC)
# the date associated with new lesion is missing, however, the date associated with RRSP1 is not (and the new lesion form should be triggered by the rrsp field


with(tu1, stopifnot(!any(duplicated(paste(USUBJID, TULNKID)))))


tu2_target <- with(subset(tu1, TUORRES %in% 'TARGET'), as.data.frame( table(USUBJID))); row.names(tu2_target) <- as.character(tu2_target$USUBJID)


tu2_nt <- with(subset(tu1, TUORRES %in% 'NON-TARGET'), as.data.frame( table(USUBJID)) ); row.names(tu2_nt) <- as.character(tu2_nt$USUBJID)


tu2_new <-  ddply(tu_new, .(USUBJID), function(d) { # get the first date when new lesion appears
  d <- d[order(d$TUDTC), ]
  d[1, ]
  
  })

stopifnot(all(sort(unique(tu_new$USUBJID)) == sort(tu2_new$USUBJID)))

stopifnot(!any(duplicated(tu2_new$USUBJID)))
row.names(tu2_new) <- as.character(tu2_new$USUBJID) 

patient_list <- unique(tu$USUBJID)

patient_tu <- data.frame( TARGET = tu2_target[patient_list,'Freq'], 
                                NT = tu2_nt[patient_list, 'Freq'],
                                NEW  = tu2_new[patient_list,'TUDTC'], stringsAsFactors = F)
row.names(patient_tu) <- patient_list


patient_tu$in_rrsp1 <- sapply(strsplit(patient_list, split='-'), function(x) tail(x, 1)) %in% with(subset(rrsp1,! RSPDN %in% 'No') , Subject)


get_tumor_list <- function(d) {
   dlply(d, .(USUBJID), function(d1) {
     unique(d1$TULNKID)
     })
}

targe_list <- get_tumor_list(subset(tu1, TUORRES %in% 'TARGET')) # list of target lesions

# check patients with more than 5 target lesions
moreThan5 <- sapply(targe_list, function(s) length(s) > 5)

if(any(moreThan5)) print( targe_list[moreThan5] )

nt_list <- get_tumor_list(subset(tu1, TUORRES %in% 'NON-TARGET'))

```

Among `r length(patient_list)` patients with tumor data, there are `r sum(is.na(patient_tu$TARGET))` patients without target lesion; there are `r sum(is.na(patient_tu$NT))` patients without non-target lesions; there are `r sum(is.na(patient_tu$NEW))` patient without new lesions.


# target lesion measurement


```{r, echo=F}

# this step is not that useful; may be skipped in future runs
tr1 <- merge(
  subset(tr, TRTEST %in% c('Longest Diameter', 'Longest Perpendicular') & TREVAL %in% c('','INVESTIGATOR' )), 
  tu1_target[, c('USUBJID','TULNKID','TUORRES','TULOC')], 
  by.x=c('USUBJID','TRLNKID'),  by.y=c('USUBJID','TULNKID'), all=T) 

# records with TUORRES is missing are for new lesions


with(subset(tr1, is.na(TUORRES)), stopifnot(all(TRGRPID %in%  'NEW')))


# in this dataset,  records with TOO SMALL TO MEASURE have no measurement; and the corresponding measurements are mapped to a separate records

tr2 <-  subset(tr1, ! ( TRORRES %in% c('ND','TOO SMALL TO MEASURE','NOT EVALUABLE') | is.na(TUORRES)))  

#=========================study specific adjustment
# 

# manually change lesion # for 30013: one of the lesion 2 should be lesion 3, there seems no way to fix this error
 
 
#==============================================


tr2 <- adjust_visit (tr2)


tr3 <- dcast(tr2 , USUBJID + TRLNKID + TULOC +   visit2 + TRDTC  ~ TRTESTCD, value.var='TRSTRESN') 

tr3$TRDTC <- as.Date(tr3$TRDTC)

# the dataset does not have both long and short diameter available
with(tr3, stopifnot( all( is.na(LDIAM)  + is.na(LPERP) >0 ))) # LDIAM: long diamater; LPERP: short diamater

table(tr3$TRLNKID) # verify that all lesions are target lesions from investigors


print('check whether nodal lesions are equivalent to LOC="LYMPH NODE"')
with(subset(tr3, !is.na(TULOC)), print( table(
  ifelse(TULOC %in%  'LYMPH NODE', 'LYMPH NODE','Other'), 
  ifelse(is.na(LDIAM), 'with short diameter','with long diameter'), useNA = 'ifany')) )
```


```{r derive SLD, echo=F}


sld <- ddply( tr3, .(USUBJID, visit2), function(d) {
  value <- with(d, ifelse( is.na(LDIAM), LPERP, LDIAM))
  lesion_cr <- with(d, ifelse( is.na(LDIAM), LPERP <10, LDIAM==0))
  d$sld <- sum( value , na.rm =F)
  d$target_cr <- all(lesion_cr)
  
  d$early_date_target <- with(subset(d, !is.na(value)), min(TRDTC), na.rm=T) # so date information is from records with valid measurement
  d$late_date_target <- with(subset(d, !is.na(value)), max(TRDTC), na.rm=T)
  
  d[1, c('USUBJID','visit2',   'target_cr', 'sld','early_date_target','late_date_target')]
  })
  


print('summary of the gap in tumor assessment day from the same visit')
with(sld, summary(as.integer(late_date_target - early_date_target)))

print('visits with missing target lesion assessment date') 
print( subset(sld, is.na(as.integer(late_date_target - early_date_target ))) )

 kable( subset(sld, late_date_target - early_date_target >5), caption = 'visits spanning more than 5 days', row.names=F)

```

```{r change date, echo=T}
# manually change early_date_target, late_date_target if necessary
# wrong dates will sort the records wrongly, and entirely mess up target response status

# sld_30154 <- subset(sld, USUBJID %in% 'BO42592-30154')
# sld_30539 <- subset(sld, USUBJID %in% 'BO42592-30539')
# 
# sld_30154[grep('TUMOR ASSESSMENT WEEK 120', sld_30154$visit2), 'early_date_target'] <- as.Date('2023-12-13') 
# sld_30154[grep('TUMOR ASSESSMENT WEEK 120', sld_30154$visit2), 'early_date_target'] <- as.Date('2023-12-13')
# 
# sld <- rbind(subset(sld, !USUBJID %in% 'BO42592-30154'), sld_30154) 


# within each subject, records are sorted by TRDTC
sld <- sld[with(sld, order(USUBJID, early_date_target, late_date_target)), ]

```
             

```{r, echo=F}
# derive change of SLD from baseline or nadir ; derive target lesion response

targetResp <- ddply(sld, .(USUBJID), function(d, baselineCondition = 'visit2 %in% "SCREENING"'){
  
  if(nrow(d) ==1) return(d)
  
  sel_baseline <- with(d, eval(parse(text = baselineCondition)))
  if(sum(sel_baseline) != 1 || !(sel_baseline[1])) {
        print(d)
    stop('cannot find one baseline SLD record')
  }
 
   
  
 d$base <-   d[ sel_baseline , 'sld']
 d$nadia <- NA
 
 for (i in  2:nrow(d) ){ # cycle through non baseline, assuming 1st line is baseline; and following lines are sorted chronocally
   prev_sld <- d[1:(i-1),'sld'] 
   d[i, 'nadia'] <- ifelse(all(is.na(prev_sld)), NA, min(prev_sld, na.rm=T))
 }
 d$chg_base <- ( d$sld - d$base ) / d$base
 d$chg_nadia <- (d$sld - d$nadia) / d$nadia
 d$abs_inc <- pmax( d$sld - d$base, d$sld - d$nadia)
 
 d$targetResp <- with(d,ifelse(is.na(chg_base), 'NE', 
                        ifelse(target_cr, 'CR', 
                        ifelse(pmax(chg_base,chg_nadia) > .2 & abs_inc >=5 ,'PD',
                        ifelse(chg_base  < -.3,'PR',  'SD')))))
 
 if(any(is.na(d$USUBJID))) {
   print(d)
   stop()
 }
 
 d
 
 
  })


with(targetResp, stopifnot(!any(is.na(early_date_target))))

```

```{r, echo=F}

with(subset(rs, RSEVAL %in% c('', 'INVESTIGATOR') & RSTEST %in% c('Non-target Response','Overall Response','Target Response')), stopifnot(!any(duplicated(paste(USUBJID, VISIT, RSTESTCD, RSDTC)))))

rs$RSDTC <- as.Date(rs$RSDTC)

 
 comp1 <- approx_merge(ds1 = subset(targetResp, ! visit2 %in% 'SCREENING') , 
  ds2 = subset(rs, RSTESTCD %in% c('TRGRESP') & RSEVAL %in% c('','INVESTIGATOR')), 
  ds1.time.var = c('early_date_target','late_date_target'), 
  ds2.time.var='RSDTC', 
  ds1.visit.var = 'visit2',
  ds2.visit.var = 'VISIT',
  ds2.vars =c('RSSTRESC') )
 
 print(lapply(comp1, dim))
  
 missed_pt_target <- subset(rs, RSTESTCD %in% c('TRGRESP') & RSEVAL %in% c('','INVESTIGATOR') & USUBJID %in% row.names(comp1$subject_not_Match)[comp1$subject_not_Match[, 2]])
 
 #with(missed_pt_target  , stopifnot(all(RSSTRESC %in% c('NA','NE'))))
 
  kable(missed_pt_target[, c('USUBJID', 'VISIT','RSSTRESC')], caption = 'patient missed in wei target response derivation', row.names=F)
      
  kable( subset(patient_tu, is.na(TARGET)), row.names =T, caption ='patients without target lesion from ttn1')
 

  
    
  kable( subset(comp1$leftover_source2, !RSSTRESC %in%  'NE', select=c('USUBJID','VISIT','RSDTC','RSTEST', 'RSSTRESC')), row.names=F, caption = 'RRSP1 records assessing target lesions that are missed by Wei')
  
  #stopifnot(all(sort(unique(comp1$matched$USUBJID)) == sort(common_subj)))

  print('response is NE or NA from wei, but there is a valid response from rrsp1')
  print( subset(comp1$matched, is.na(gap_day)  & ! RSSTRESC %in% c(NA,'NA','NE'))) # when gap_day is missing, the most common reason is that targetResp (wei derivation) is missing
  
  
  
  kable( subset(comp1$matched, gap_day >=40, select=c('USUBJID', 'visit2','early_date_target','late_date_target','RSDTC','gap_day')), row.names=F, caption ='dates from target lesion assessment and recist assessment are 40 days or more apart')
  
with(comp1$matched, table(RSSTRESC, targetResp, useNA = 'ifany')) 

diff_target <- subset(comp1$matched, ( is.na(RSSTRESC) != is.na(targetResp) ) | ( !is.na(RSSTRESC) &  !is.na(targetResp) & RSSTRESC != targetResp ) ) 

write.csv(diff_target[, c('USUBJID','visit2','RSSTRESC', 'sld','early_date_target',	'late_date_target',	'base',	'nadia',	'chg_base',	'chg_nadia',	'targetResp')]
, file = paste(outF,'diff_target.csv', sep = '/'), row.names=F)

```

# non-target measurement

```{r, echo=F}
 tr4_check <- merge(
   subset(tr, TRTEST %in% c('Tumor State') & TREVAL %in% c('', 'INVESTIGATOR') &  TRGRPID %in% 'NON-TARGET'), 
   tu1_nt[, c('USUBJID','TULNKID','TUORRES','TULOC')], 
   by.x=c('USUBJID','TRLNKID'),  by.y=c('USUBJID','TULNKID'), all.x=T) # only left join, many non-target lesion records are not populated


# all NON-PATHOLOGICAL should be from lymph node
with(tr4_check, table(ifelse(TULOC %in% "LYMPH NODE", 'nodal','non-nodal'),
                      ifelse(TRSTRESC %in% c('NON-PATHOLOGICAL'), 'node_cr','other'), useNA = 'ifany'))

with(tr4_check, table(ifelse(TULOC %in% "LYMPH NODE", 'nodal','non-nodal'),
                      ifelse(TRSTRESC %in% c('ABSENT'), 'non-node_cr','other'), useNA = 'ifany'))


tr4 <- subset(tr, TRTEST %in% c('Tumor State') & TREVAL %in% c('','INVESTIGATOR') &  TRGRPID %in% 'NON-TARGET') 



tr4b <- adjust_visit_nt (tr4,tlist = nt_list )

tr4b$TRDTC <- as.Date(tr4b$TRDTC)

stopifnot(!any(is.na(tr4b$TRDTC))) 


ntRespByVisit <- ddply(tr4b, .(USUBJID, visit2), function(d1, tlist = nt_list, ne_code = c(NA,'NOT ASSESSED')){
  s <- d1[1,'USUBJID']
  stopifnot(! s %in% tlist)
  
  d1$ntRespVisit <- with(d1, ifelse( any(TRSTRESC %in% 'UNEQUIVOCAL PROGRESSION'), 'PD',
                        ifelse( any(TRSTRESC %in% ne_code) || nrow(d1) < length( d1$TRLNKID), 'NE',
                        ifelse( all(TRSTRESC %in% c('NON-PATHOLOGICAL','ABSENT')),'CR', 'NON-CR/NON-PD'))))

  d1$early_date_nt <- with(subset(d1, !TRSTRESC %in% ne_code), min(TRDTC), na.rm=T)  
  d1$late_date_nt <- with(subset(d1, !TRSTRESC %in% ne_code), max(TRDTC), na.rm=T)
    
  d1[1,  c('USUBJID','visit2',   'ntRespVisit','early_date_nt','late_date_nt')] 
  })

#subset(ntRespByVisit, is.infinite(early_date_target))

# the ddply function sorts records by the value of visit2; I need sort it by date
ntRespByVisit  <- ntRespByVisit[with(ntRespByVisit , order(USUBJID, early_date_nt)), ]

ntResp <- ddply(ntRespByVisit, .(USUBJID), function(d){
  if( any (d$ntRespVisit %in% 'CR')) { 
  early_cr <- min(which(d$ntRespVisit %in% 'CR'))
  for (i in 2:nrow(d)) {
    if(i > early_cr & d[i,'ntRespVisit'] %in% 'NON-CR/NON-PD') d[i,'ntRespVisit'] <- 'PD'
  }
  }
  d
  })

comp2 <- approx_merge (
  ds1 = subset( ntResp, ! visit2 %in% 'SCREENING') , 
  ds2 = subset(rs, RSTESTCD %in% c('NTRGRESP') & RSEVAL %in% c('','INVESTIGATOR')), 
  ds1.time.var = c('early_date_nt','late_date_nt'), 
  ds2.time.var='RSDTC', 
  ds1.visit.var = 'visit2',
  ds2.visit.var = 'VISIT',
  ds2.vars =c('RSSTRESC')) 

print(lapply(comp2, dim))
#kable( subset(comp2$matched, USUBJID %in% 'BO42592-30265'), row.names=F, caption ='all non target assessment for BO42592-30265 in the matched non-target lesion data')

#======= check comparison

 missed_pt_nt <- subset(rs, RSTESTCD %in% c('NTRGRESP') & RSEVAL %in% c('','INVESTIGATOR') & USUBJID %in% row.names(comp2$subject_not_Match)[comp2$subject_not_Match[, 2]])
 
 if( with(missed_pt_nt  , !all(RSSTRESC %in% c('NA','NE')))) {
   print ('patients without non-target lesion assessment')
   print(subset(missed_pt_nt, !RSSTRESC %in% c('NA','NE'), select=c('USUBJID','VISIT','RSTESTCD','RSSTRESC')))
   
   
 }else{
  stopifnot(all( missed_pt_nt$USUBJID %in% setdiff(patient_list, names(nt_list))))  # so these patients do not have non-target lesions
 } 
```

```{r, echo=F}
 
print('RRSP1 records assessing non-target leions that are missed by Wei')
  print( subset(comp2$leftover_source2, !RSSTRESC %in%  'NE', select=c('USUBJID','VISIT','RSDTC','RSSTRESC'))  )
  
  #stopifnot(all(sort(unique(comp2$matched$USUBJID)) == sort(common_subj)))

  print('response is NE or NA from wei, but there is a valid response from rrsp1')
  print( subset(comp2$matched, is.na(gap_day)  & ! RSSTRESC %in% c(NA,'NA','NE'))) # when gap_day is missing, the most common reason is that targetResp (wei derivation) is missing
  

  kable( subset(comp2$matched, gap_day >=40, select=c('USUBJID', 'visit2','early_date_nt','late_date_nt','RSDTC', 'gap_day')), row.names=F, caption ='dates from non-target lesion assessment and recist assessment are 40 days or more apart')
  

#======

with(comp2$matched, table(RSSTRESC, ntRespVisit, useNA = 'ifany'))


diff_nt <- subset(comp2$matched, RSSTRESC != ntRespVisit  )

write.csv(diff_nt[, c('USUBJID','visit2','RSSTRESC', 'early_date_nt',	'late_date_nt',	'ntRespVisit')]
, file = paste(outF,'diff_nontarget.csv', sep= '/'), row.names=F)

```

# new lesion

```{r, echo=F}

print('new lesions from RRSP1 and new lesion form are mapped together')

with( subset(tu, grepl('NEW', TUORRES) & TUEVAL %in% c('', 'INVESTIGATOR')), table(substring(TUSPID, 1,10)))

if(F){
  # new lesion from the new lesion assessment form
new1 <- unique( subset(tu, grepl('NEW', TUORRES) & grepl('IRNLES', TUSPID), select=c('USUBJID', 'VISIT','TUDTC')) )

kable(subset(new1, paste(USUBJID, VISIT ) %in% paste(USUBJID, VISIT )[duplicated(paste(USUBJID, VISIT ))]), row.names=F, caption ='Tumor Assessment New Lesions: duplicated USUBJID-Visit') 

new1 <- new1[with(new1, !duplicated(paste(USUBJID, VISIT))), ]

new2 <-  unique( subset(tu, grepl('NEW', TUORRES) & grepl('RRSP1', TUSPID), select=c('USUBJID', 'VISIT','TUDTC','TULNKID')) )
new2$TUDTC <- as.Date(new2$TUDTC)

}
stopifnot(all( grepl('RRSP1', tu_new$TUSPID)))
new2 <-  tu_new[,  c('USUBJID', 'VISIT','TUDTC','TULNKID')]


with(new2, stopifnot(!any(duplicated(paste(USUBJID, VISIT, TUDTC)))))

```

# overall response

rederive overall response based on target / non target / new lesion status from the RECIST 1.1 form

```{r, echo=F}


resp1 <-   dcast(subset(rs,  RSEVAL %in% c('', 'INVESTIGATOR') & RSTESTCD %in% c('TRGRESP','NTRGRESP','OVRLRESP')), USUBJID+ VISIT + RSDTC ~ RSTESTCD, value.var='RSSTRESC')

resp1$visit2 <- with(resp1, ifelse(VISIT %in% "UNSCHEDULED", paste(VISIT, RSDTC), VISIT))
resp1$VISIT <- NULL
resp1$RSDTC2 <- resp1$RSDTC


#stopifnot( all(sort(unique(resp1$USUBJID)) == sort( row.names(patient_tu[patient_tu$in_rrsp1, ]))) )

# patients without new lesion
resp2a <- subset(resp1, ! USUBJID %in% new2$USUBJID)

#stopifnot( all(sort(unique( resp2a$USUBJID)) == sort( row.names(subset(patient_tu, is.na(NEW) & in_rrsp1))) ))
 

# patients with new lesion
resp2b <- approx_merge (
  ds1 = subset(resp1, USUBJID %in% new2$USUBJID) , 
  ds2 = new2, 
  ds1.time.var = c('RSDTC','RSDTC2'), 
  ds2.time.var='TUDTC', 
  ds1.visit.var = 'visit2',
  ds2.visit.var = 'VISIT',
  ds2.vars =c('TULNKID')) 

stopifnot(nrow(resp2b$leftover_source2) == 0)
stopifnot(nrow(resp2b$subject_not_Match) == 0)


stopifnot(all(sort(unique(resp2b$matched$USUBJID)) == sort( row.names(subset(patient_tu, !is.na(NEW) & in_rrsp1))) ))

colnames(resp2b$matched) <- gsub('TULNKID','NEW', colnames(resp2b$matched))
 
resp2 <-  rbind.fill(resp2b$matched, resp2a)

resp2$NEW <- ifelse(grepl('NEW', resp2$NEW),'Yes','No')

 
  
  resp2[ ,'overall'] <- ifelse( !is.na( patient_tu[resp2$USUBJID,'TARGET'] ), 
                                        
       with(resp2, case_when( # with target lesion
          TRGRESP  %in% "PD" | NTRGRESP %in% "PD" | NEW %in% "Yes" ~ "PD",
          TRGRESP %in% "CR" & NTRGRESP %in% "CR" & NEW %in% "No" ~ "CR",
          TRGRESP %in% "CR" & NTRGRESP %in% "NA" & NEW %in% "No" ~ "CR",
          TRGRESP %in% "CR" & toupper(NTRGRESP) %in% c("NON-CR/NON-PD", "NE") & NEW %in% "No" ~ "PR",
          TRGRESP %in% "PR" & toupper(NTRGRESP) %in% c("CR", "NON-CR/NON-PD", "NE", "NA") & NEW %in% "No" ~ "PR",
          TRGRESP %in% "SD" & toupper(NTRGRESP) %in% c("CR", "NON-CR/NON-PD", "NE", "NA") & NEW %in% "No" ~ "SD",
          TRGRESP %in% "NE" & toupper(NTRGRESP) %in% c("CR", "NON-CR/NON-PD", "NE", "NA") & NEW %in% "No" ~ "NE",
          TRGRESP %in% "NA" & toupper(NTRGRESP) %in% "NON-CR/NON-PD" & NEW %in% "No" ~ "NON-CR/NON-PD",
          TRGRESP %in% "NA" & NTRGRESP %in% "CR" & NEW %in% "No" ~ "CR",
          TRGRESP %in% "NA" & NTRGRESP %in% c("NE", "NA") & NEW %in% "No" ~ "NE",
          TRUE ~ "NA"
        )),
       
      with(resp2, case_when( # no target lesion
          NTRGRESP %in% "PD" | NEW %in% "Yes" ~ "PD",
          NTRGRESP %in% "CR" & NEW %in% "No" ~ "CR",
          NTRGRESP %in% "NON-CR/NON-PD" & NEW %in% "No" ~ "NON-CR/NON-PD",
          NTRGRESP %in% c("NE", "NA") & NEW %in% "No" ~ "NE",
          TRUE ~ "NA"
        ) )
  )
  

  
  with(resp2, print( table( overall, OVRLRESP, useNA ='ifany')) )
  
  write.csv( subset(resp2, overall!= OVRLRESP), file =   paste(outF, 'diff_overall.csv', sep = '/'), row.names=F)
  
```

