---
title: "simulation to study strategies to handle small strata"
author: "Wei Zou"
date: "`r Sys.time()`"
output: pdf_document
---

Version history

* sim_strata_1b: use simsurv to replace rpact as the engine to simulate survival data. The simulation cannot handle ramp up enrollment or dropoff; but it can simulate data with high EPR easily

* sim_strata_1: code is modified from C:/Users/zouw2/aPDL1/impower010/natera/programs/sim_collapsibility4_test_strata


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)




library(knitr)

library(ggplot2)

 source('~/R/lib_2020/data_functions.r')
source('~/R/lib_2020/output_functions.r')

spec <- list(home  = '~/projects/stratification' )

```

 

```{r send jobs, eval=F}

# this chuck is supposed to be run separately

# run this first: #  ml load R

prep_cmd <- function(i, s = spec){
  
  log_home <- paste(spec$home, 'log', spec$run_name, sep = '/')
  if(!file.exists(log_home))   dir.create(log_home, recursive =T)
  
  log_out <-  paste(log_home, '/', i,'.cmd', sep='')
  cat("#!/bin/bash
    #SBATCH -J strat_sim
    #SBATCH --qos=short
    #SBATCH --cpus-per-task 4
    #SBATCH --mem-per-cpu=4G
    #SBATCH -o sim.o%J        
    #SBATCH -e sim.e%J
    ml R/R_4.1.0_Bioc_3.13/R-4.1.0-Bioc-3.13-prd-20210913
   
R CMD BATCH --no-save --no-restore '--args ", i+s$seed.base, " ", s$n, " ", s$dup, ' ', s$hr,' ', s$med,' ', s$readoutTime,' ',s$run_name, "' ",  s$home, "/", s$run_code,' ', log_home,"/run_",i,".log "
    ,
      file =log_out, sep='')
  
  return(log_out)
  
}


M <- 100

spec <- list(home  = '~/projects/stratification', run_name =  '1F_scen1b',
             seed.base= 27518, n = '60,60',  dup = '20', hr = '0.8,0.8', med = '27.7,17.2', readoutTime= '50', run_code = 'strat1b.r' , total_run=M)

system(paste('cd', paste(spec$home,'log', sep ='/')))

#for (j in seq(1, spec$total_run , by= spec$n_run)) {
for (j in seq(1, spec$total_run , by= 1)) {
  system(paste('sbatch', prep_cmd(j)))
}


spec$n <- '12,108' ; spec$run_name <- '1F_scen2b'
for (j in seq(1, spec$total_run , by= 1)) {
  system(paste('sbatch', prep_cmd(j)))
}

spec$n <- '60,60'; spec$med <- c('27.7,11'); spec$run_name <- '1F_scen3b'; spec$readoutTime <- 40

for (j in seq(1, spec$total_run , by= 1)) {
  system(paste('sbatch', prep_cmd(j)))
}

spec$n <- '12,108'; spec$med <- c('27.7,11'); spec$run_name <- '1F_scen4b'

for (j in seq(1, spec$total_run , by= 1)) {
  system(paste('sbatch', prep_cmd(j)))
}

```


```{r additional runs, eval =F}

spec <- list(home  = '~/projects/stratification', run_name =  '1F_scen2c',
             seed.base= 27518, n = '12,108',  dup = '20', hr = '0.8,0.8', med = '27.7,17.2', readoutTime= '100', run_code = 'strat1b.r' , total_run=M)

system(paste('cd', paste(spec$home,'log', sep ='/')))

 
for (j in seq(1, spec$total_run , by= 1)) {
  system(paste('sbatch', prep_cmd(j)))
}
```

# prevalence 0.5 , prognostic HR .63
 
```{r, warning=FALSE }

 r1 <- readFolder.rdata( paste(spec$home,'reports/1F_scen1b'  , sep = '/'), objName = 'r1')

knitr::kable( summarizeMatrix(r1[, c('n','nevent', 'p_s','delta_p','O-E_M2','O-E_M1')]), digits=3)

  
```

# prevalence 0.1 , prognostic HR .63

```{r, warning=FALSE }

 
  
r2 <-   readFolder.rdata( paste(spec$home,'reports/1F_scen2b'  , sep = '/'), objName = 'r1')
 
knitr::kable( summarizeMatrix(r2[, c('n','nevent','p_s','delta_p','O-E_M2','O-E_M1')]), digits = 3)
 
  
``` 
 
## more mature read at 100 month
```{r, warning=FALSE }

 
  
r2c <-   readFolder.rdata( paste(spec$home,'reports/1F_scen2c'  , sep = '/'), objName = 'r1')
 
knitr::kable( summarizeMatrix(r2c[, c('n','nevent','p_s','delta_p','O-E_M2','O-E_M1')]), digits = 3)
 
  
``` 

 
# prevalence 0.5 , prognostic HR .4
 
```{r, warning=FALSE}

 r1 <-   readFolder.rdata( paste(spec$home,'reports/1F_scen3b'  , sep = '/'), objName = 'r1')

 
knitr::kable( summarizeMatrix(r1[, c('n','nevent', 'p_s','delta_p','O-E_M2','O-E_M1')]), digits=3)

  
```
 

# prevalence 0.1 , prognostic HR .4

```{r, warning=FALSE}

r2 <-   readFolder.rdata( paste(spec$home,'reports/1F_scen4b'  , sep = '/'), objName = 'r1')
 
knitr::kable( summarizeMatrix(r2[, c('n','nevent','p_s','delta_p','O-E_M2','O-E_M1')]), digits = 3)
 
  
``` 